Co-moving stars in \textsl{Gaia DR1}:\\ An abundance of very wide separation co-moving pairs

### 2016-12-02 paper/ms.tex
more text edits

\begin{abstract}
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
presents an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
Here, we look for high-confidence co-moving pairs of stars using this data.
We use a justified marginalized likelihood ratio test to separate
pairs of stars with similar 3D velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
We find 13,085 pairs of co-moving stars among 10,606 unique stars
out to very wide separations up to 10~pc, the limit of our search.
Some of these pairs form larger groups through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
We also report many new co-moving groups of pairs.
More surprisingly, we find a large number
of very wide ($>3$~pc) separation co-moving pairs of stars.
The number of these very wide separation pairs increase with increasing
separation; and cannot be explained purely by false-positives contamination.
We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
with auxiliary information, and discuss its potential use
for making dynamical inferences, and testing stellar models and chemical
abundance measurements.
\end{abstract}

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence co-moving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$ that we describe later.
\end{itemize}

Under these assumptions, the likelihood of a proper motion measurement,
$\bs{\mu}$, for a star with true distance, $d$, and true tangential velocity,
$\vec{v}_t =
(\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
\begin{align}
  L(\vec{\mu} \given \vec{v}, d) &=
    \left[\det\left(\frac{{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\vec{\mu} - \vec{x}_\theta \right)} \,
    {\mat{C}}^{-1} \,
    \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
  \vec{x}_\theta &= d^{-1} \, \vec{v}_t
\end{align}
where the tangential velocity $\vec{v}_t$ is related to the 3D velocity
$\vec{v}$ by projection matrix $\mat{M}$ at the star's sky position
$(\alpha, \delta)$
\begin{align}
  \vec{v}_t &= \mat{M}\,\vec{v} \\
  & = \left(
      \begin{array}{c c c}
        -\sin\alpha & \cos\alpha & 0 \\
        -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta
      \end{array}
    \right) \,
    \left(\begin{array}{c} v_x \\ v_y \\ v_z \end{array}\right) \label{eq:transformation}
\end{align}
%and the modified covariance matrix $\tilde{\mat{C}}$ is
%\begin{equation}
%  \tilde{\mat{C}} = \mat{C} + s^2 \, \eye
%\end{equation}

For a given pair, we compute the fully marginalized likelihood (FML) for the hypotheses (1) and (2),
$\mathcal{L}_1$ and $\mathcal{L}_2$.
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
candidate co-moving pairs, as described in $\sectionname~\ref{sub:selection}$ in
more detail.
To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
L_j$, are marginalized over the (unknown) true 3D velocity and distance for each star in
the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
    L_i(\vec{v}, d_i, s^2) \,
    L_j(\vec{v}, d_j, s^2) \,
    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
  \mathcal{L}_2 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{v}_i, d_i, 0) \,
    L_j(\vec{v}_j, d_j, 0) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
\end{align}
Here, $p(d_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
$\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
Note that the FML for the hypothesis (1) involves integration over one velocity $\vec{v}$ that
generates the likelihoods for both stars, $L_i$ and $L_j$.
The marginalization integral for hypothesis (2) can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
\begin{equation}
  Q_i = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi_i)
\end{equation}

@@

For wide binaries, the assumption of the stars having the same 3D velocity for 
hypothesis (1) can
break down for high precision proper motion measurements because of the orbital
velocity (blue solid line in Figure~\ref{fig:dv-sep}).
To account for this, we modify set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.

@@

In this section, we examine the distribution of (log-)likelihood ratios
$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
for this quantity to select co-moving pairs from the initial pairs sample.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
    The black dots (upper right) and lines (lower left) correspond to the entire initial sample of pairs.
    For comparision, we also present the likelihood ratio distribution of random pairs
    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
    candidate co-moving pairs.
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k initial pairs examined.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of the likelihood ratios of pairs
in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. This is a cylindrical
    projection of onto the Galactic plane. The angle is the Galactic
    longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
    of distance using Equation~\ref{eq:dist}.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as light blue circles,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    gray circles.
    Some of the larger connected components are clearly associated with known
    clusters, but we also discover quite a few new co-moving groups.
    \label{fig:glon_d_pairlines}}
\end{figure*}

@@

However, there are still many new larger connected components that we
discover.
If we define a condition to associate a connected component to a known cluster
as having more than 3 members within 10~pc of the nominal position of
the cluster, for the 61 connected components with size larger than 5,
we find that only 10 are associated with a cluster cataloged in
\citet{Kharchenko:2016aa}.
It is also worth noting that the positions of some of the known clusters
are offset from those of the connected components associated with them,
indicating that the \tgas\ data improves the distance estimates of these clusters.
Finally, not all known star clusters are recovered in our search.
This is primarily because of the non-uniform coverage of \tgas\ data.

Ultimately, any candidate co-moving pair found in this work
needs to be verified using radial velocities.
Here, we use 210,368 cross-matches of the \tgas\ with
the Radial Velocity Experiment (\rave) to
assess the false-positive rates of our selection.
We have 283 pairs with both stars matched with \rave.
Figure~\ref{fig:raverv} shows the difference in radial velocities
between the two stars in a pair, $\Delta v_r$, as a function of their physical separation.
We show $\Delta v_r$ in units of $\sigma_{\Delta v_r}$ which we estimate
as the quadrature sum of $\sigma _{v_r}$ for each star.
The fraction of pairs with good agreement in radial velocity decreases with
increasing separation.
This, after all, is not surprising because we are only using
2D velocity information (proper motions) with errors.
However, the contamination becomes significant only at $>1$~pc
(which depends on the local stellar number density and velocity dispersion).
Given the excellent correspondence between
aggregates of co-moving pairs (connected components)
and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
we may expect that pairs in these larger connected components,
which will often have separations $>1$~pc,
to have less contamination from false-positives.
We divide the co-moving pairs into those
mutually exclusively connected (i.e., in a connected component of
size 2), and those in a larger group.
We indeed find that many pairs in larger groups are at $>1$~pc, yet
the fraction of pairs that have identical radial velocities within $3\sigma$ is
higher than the mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.

@@

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.
As expected, pairs in larger connected components are mostly
found with separations larger than 1~pc.
Surprisingly, however, we also find a large number of mutually exclusive co-moving
pairs
at $>1$~pc as well. Even if we consider the increasing false-positive rate at large
separations, the distribution is not significantly changed as the number of pairs
at $>1$~pc is in fact increasing much faster (as a power-law)
than the decrease due to the false-positives
(bottom panel of Figure~\ref{fig:raverv}).
The nature of these very large separation, mutually exclusive pairs,
which cannot be gravitationally bound to each other, needs further investigation.
Can they be remnants of escaped binaries that are drifting apart?
In a study of the evolution of wide binaries including the Galactic tidal field
as well as passing field stars, \citet{Jiang:2010aa} found that we expect
to find a peak at $\sim 100-300$~pc in the projected separation due to
stars that were once in a wide binary system, but are drifting apart with small
relative velocities ($\sim 0.1$~\kms).
Our search which stops on the maximum radius of 10~pc is obviously insufficient
to clearly detect this peak, but it shows such hint, and demonstrates
that such large scale phase-space correlations may now be looked for and studied within the \gaia\ data.

@@

Finally, we present the color-magnitude diagrams of co-moving pairs
using the cross-matches with \tmass.
A more detailed study of stellar parameters using photometry from various sources
will follow. Figure~\ref{fig:cmd_large} and \ref{fig:cmd_me} shows $G-J$ vs $G$
color-magnitude diagrams for stars in larger connected components (size$>2$)
and in mutually exclusive pairs, respectively.
The connected components shown in Figure~\ref{fig:cmd_large}
correspond to those visualized
in Figure~\ref{fig:graphviz_examples}.
For stars in larger co-moving groups, there is a noticeable
lack of evolved, off-main sequence
stars, in agreement with these kinematic structures being young.
For mutually exclusive pairs, we divide the pairs by separation at 1~pc
above which the false-positive rate due to random pairs starts to increase.
While many pairs are located along the main sequence, we also find
quite a few of main sequence-red giant pairs,
which will be valuable to anchoring stellar atmospheric models together.
%TODO: SMOH: what else can i discuss?

@@

In this section, we describe our catalog of candidate co-moving pairs of stars.
For 13,058 pairs (10,606 unique stars) found in this work,
we provide the following information:
%
\begin{itemize}
  \item \tgas\ source id for the two stars in a pair; these may be used to easily
    retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and
    \rave\ using the \gaia\ data archive.
  \item physical separation between the two stars in pc
  \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
  \item size of the connected component that the pair belongs to
  \item unique ID of the connected component that the pair belongs to; we assigned
    a unique ID in descending order of size, thus ID~0 is the largest connected component with 151 stars that corresponds to the Pleiades.
\end{itemize}
%
Additionally, given the strong correlation between the larger connected components
of co-moving pairs
and the known star clusters and OB associations, we provide for each star
%
\begin{itemize}
  \item index of the closest cluster in the Milky Way Star Cluster catalog of
  \citet{Kharchenko:2016aa} using 3D separation
  \item distance to the closest cluster in pc
  \item the ID of the OB association (A-L) if the star is in the catalog of
  stars in the known OB associations by \citet{de-Zeeuw:1999aa}
\end{itemize}
%
A part of the entire table is presented in Table~\ref{tab:catalog}.

@@

In this \documentname, we looked for co-moving pairs of stars in the
\tgas\ data released as part of the \gaia\ DR1.
Our method is to compare
the fully marginalized likelihoods between the two hypotheses, that a pair of
stars shares the same 3D velocity, and that the two stars have two independent
3D velocities, incorporating the covariances of parallax and proper motions.
We argued for a reasonable cut of the likelihood ratio, and found
13,058 candidate pairs of co-moving stars among 10,606 unique stars
with separation ranging from 0.005~pc to 10~pc, the limit of our search.

@@

We have also found a surprisingly large number of very wide separation ($>3$~pc)
mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
of each other, and not part of large connected components.
These are most likely remnants of dissolving wide binaries and
open clusters. If so, this population should still be relatively young compared
to the general disk field population. Modeling the color-magnitude diagram
distribution of these stars can shed some light on this issue.
If they are truly dissolving stars that were born coeval,
the sample of very wide separation co-moving pairs can potentially be used
to measure the recent ($\lesssim 1$~Gyr)
star formation history in the Solar neighborhood.
Co-moving stars with separation less than 1~pc are very promising candidates
for wide binaries. They are found to be pairs of stars of varying stellar types.
Some of these pairs, such as main sequence-red giant or F/G/K-M dwarfs,
will be particularly valuable for testing theoretical
stellar models and calibrating observational measurements of low mass stars.

We note that a similar search for wide binaries using the \tgas\ data is performed
in a recent work by \citet{2016arXiv161107883O}. They find $\approx 1900$ wide binaries
with separation typically less than 1.5~pc, and 256 pairs with separation larger than
$\sim 1$~pc.
We emphasize that our method is based on a probabilistic model for the assumptions
on the \emph{3D velocities} of the two stars in a pair, and that we
marginalize over the (unknown) three components of the velocities 
in contrast to just working in the proper motion space.

>>>>

\begin{abstract}
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
presents an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
Here, we look for high-confidence co-moving pairs of stars using this data.
We use a justified marginalized likelihood ratio test to separate
pairs of stars with similar 3D velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using
(bias-corrected) inverse parallax as point-estimate of distance,
the likelihood ratio test
is based on a probabilistic model that includes the covariances
of parallax and proper motions, and marginalizes the unknown
distances and 3D velocities of stars.
We find 13,085 pairs of co-moving stars among 10,606 unique stars
out to very wide separations up to 10~pc, the limit of our search.
Some of these pairs form larger groups through mutual co-moving neighbors,
many of which correspond to the known open clusters, and OB associations.
We also report many new co-moving groups of pairs.
More surprisingly, we find a large number
of very wide ($>3$~pc) separation co-moving pairs of stars.
The number of these very wide separation pairs increase with increasing
separation; and cannot be explained purely by false-positive contamination.
We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
with auxiliary information, and discuss its potential use
for making dynamical inferences, and testing stellar models and chemical
abundance measurements.
\end{abstract}

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence co-moving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$.
\end{itemize}

Under these assumptions, the likelihood of a proper motion measurement,
$\bs{\mu}$, for a star with true distance, $d$, and true 3D velocity $\vec{v}$ is
\begin{align}
  L(\vec{\mu} \given \vec{v}, d, s^2) &=
    \left[\det\left(\frac{\tilde{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\vec{\mu} - \vec{x}_\theta \right)} \,
    \tilde{\mat{C}}^{-1} \,
    \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
  \vec{x}_\theta &= d^{-1} \, \vec{v}_t
\end{align}
where the tangential velocity $\vec{v}_t = (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$
is related to the 3D velocity
$\vec{v}$ by projection matrix $\mat{M}$ at the star's sky position
$(\alpha, \delta)$
\begin{align}
  \vec{v}_t &= \mat{M}\,\vec{v} \\
  & = \left(
      \begin{array}{c c c}
        -\sin\alpha & \cos\alpha & 0 \\
        -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta
      \end{array}
    \right) \,
    \left(\begin{array}{c} v_x \\ v_y \\ v_z \end{array}\right) \label{eq:transformation}
\end{align}
and the modified covariance matrix $\tilde{\mat{C}}$ is
\begin{equation}
  \tilde{\mat{C}} = \mat{C} + (s/d)^2 \, \eye
\end{equation}
The parameter $s$ is added to allow for small tolerance in velocities which we discuss below.

For a given pair, we compute the fully marginalized likelihood (FML) for the hypotheses (1) and (2),
$\mathcal{L}_1$ and $\mathcal{L}_2$.
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
candidate co-moving pairs, as described in $\sectionname~\ref{sub:selection}$ in
more detail.
To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
L_j$, are marginalized over the (unknown) true distance and 3D velocity for each star in
the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
    L_i(\vec{\mu_i} \given \vec{v}, d_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}, d_j, s^2) \,
    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
  \mathcal{L}_2 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{\mu_i} \given \vec{v}_i, d_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}_j, d_j, s^2) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
\end{align}
Here, $p(d_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
$\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
Note that the FML for the hypothesis (1) involves integration over one velocity $\vec{v}$ that
generates the likelihoods for both stars, $L_i$ and $L_j$.
The marginalization integral for hypothesis (2) can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q(\vec{\mu_i}, \varpi_i) \, Q(\vec{\mu_j}, \varpi_j)$ where
\begin{equation}
  Q(\vec{\mu}, \varpi) = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, d, s^2) \,
  p(\vec{v}) \, p(d \given \varpi)
\end{equation}

@@

For wide binaries, the assumption of the stars having the same 3D velocity for 
hypothesis (1) can
break down for high precision proper motion measurements because of the orbital
velocity (blue solid line in Figure~\ref{fig:dv-sep}).
To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
Because $s^2$ is much smaller than the velocity dispersions of the velocity prior
$p(\vec{v})$, it has minimal effect on the hypothesis (2).

@@

In this section, we examine the distribution of (log-)likelihood ratios
$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
for this quantity to select co-moving pairs from the initial sample.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
    The black dots in the upper right panel and
    the black line in the lower lef panel correspond to the entire initial sample of pairs.
    For comparision, we also present the likelihood ratio distribution of random pairs
    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
    candidate co-moving pairs.
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k pairs in the initial sample.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between the likelihood
ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of the likelihood ratios of pairs
in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. This is a cylindrical
    projection of onto the Galactic plane. The angle is the Galactic
    longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
    of distance using \eqname~\ref{eq:dist}.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as light blue circles,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    gray circles.
    Some of the larger connected components are clearly associated with known
    clusters, but we also discover quite a few new co-moving groups.
    \label{fig:glon_d_pairlines}}
\end{figure*}

@@

However, there are still many new larger connected components that we
discover.
If we define a condition to associate a connected component to a known cluster
as having more than 3 members within 10~pc from the nominal position of
the cluster, for the 61 connected components with sizes larger than 5,
we find that only 10 are associated with a cluster cataloged in
\citet{Kharchenko:2016aa}.
It is also worth noting that the positions of some of the known clusters
are offset from those of the connected components associated with them,
indicating that the \tgas\ data improves the distance estimates of these clusters.
Finally, not all known star clusters are recovered in our search.
This is primarily because of the non-uniform coverage of the \tgas\ data.

Ultimately, any candidate co-moving pair found in this work
needs to be verified using radial velocities.
Here, we use 210,368 cross-matches of the \tgas\ with
Radial Velocity Experiment (\rave) to
assess the false-positive rates of our selection.
We have 283 pairs with both stars matched with \rave.
Figure~\ref{fig:raverv} shows the difference in radial velocities
between the two stars in a pair, $\Delta v_r$, as a function of their physical separation.
We show $\Delta v_r$ in units of $\sigma_{\Delta v_r}$ which we estimate
as the quadrature sum of $\sigma _{v_r}$ for each star.
The fraction of pairs with good agreement in radial velocity decreases with
increasing separation.
This, after all, is not surprising because we are only using
2D velocity information (proper motions) with errors.
However, the contamination becomes significant only at $>1$~pc
(which depends on the local stellar number density and velocity dispersion).
Given the excellent correspondence between
aggregates of co-moving pairs (connected components)
and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
we may expect that pairs in these larger connected components,
which will often have separations $>1$~pc,
to have less contamination from false-positives.
We divide the co-moving pairs into those
mutually exclusively connected (i.e., in a connected component of
size 2), and those in a larger group.
We indeed find that many pairs in larger groups are at $>1$~pc, yet
the fraction of pairs that have identical radial velocities within $3\sigma$ is
higher than the mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.

@@

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.
As expected, pairs in larger connected components are mostly
found with separations larger than 1~pc.
Surprisingly, however, we also find a large number of mutually exclusive co-moving
pairs
at $>1$~pc as well. Even if we consider the increasing false-positive rate at large
separations, the distribution is not significantly changed as the number of pairs
at $>1$~pc is in fact increasing much faster (as a power-law)
than the decrease due to the false-positives
(bottom panel of Figure~\ref{fig:raverv}).
The nature of these very wide separation, mutually exclusive pairs,
which cannot be gravitationally bound to each other, needs further investigation.
Can they be remnants of escaped binaries that are drifting apart?
In a study of the evolution of wide binaries including the Galactic tidal field
as well as passing field stars, \citet{Jiang:2010aa} found that we expect
to find a peak at $\sim 100-300$~pc in the projected separation due to
stars that were once in a wide binary system, but are drifting apart with small
relative velocities ($\sim 0.1$~\kms).
Our search which stops on the maximum radius of 10~pc is obviously insufficient
to clearly detect this peak, but it shows such hint, and demonstrates
that such large scale phase-space correlations may now be looked for and studied using the \gaia\ data.

@@

Finally, we present the color-magnitude diagrams of co-moving pairs
using the cross-matches with \tmass.
A more detailed study of stellar parameters using photometry from various sources
will follow. Figure~\ref{fig:cmd_large} and \ref{fig:cmd_me} shows $G-J$ vs $G$
color-magnitude diagrams for stars in larger connected components (size$>2$)
and in mutually exclusive pairs, respectively.
The connected components shown in Figure~\ref{fig:cmd_large}
correspond to those visualized
in Figure~\ref{fig:graphviz_examples}.
For stars in larger co-moving groups, there is a noticeable
lack of evolved, off-main sequence
stars, in agreement with these kinematic structures being young.
For mutually exclusive pairs, we divide the pairs by separation at 1~pc
above which the false-positive rate due to random pairs starts to increase.
While many pairs are located along the main sequence, we also find
quite a few of main sequence-red giant pairs,
which will be valuable to anchoring stellar atmospheric models together.

@@

In this section, we describe our catalog of candidate co-moving pairs of stars.
For 13,058 pairs (10,606 unique stars) found in this work,
we provide the following information:
%
\begin{itemize}
  \item \tgas\ source id for the two stars in a pair; these may be used to easily
    retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and
    \rave\ using the \gaia\ data archive.
  \item physical separation between the two stars in pc. Note that this is calculated
    using the point-estimates of distances (\eqname~\ref{eq:dist}) of the two stars.
  \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
  \item size of the connected component that the pair belongs to
  \item unique ID of the connected component that the pair belongs to; we assigned
    a unique ID in descending order of size, thus ID~0 is the largest connected component with 151 stars that corresponds to the Pleiades.
\end{itemize}
%
Additionally, given the strong correlation between the larger connected components
of co-moving pairs
and the known star clusters and OB associations, we provide for each star
%
\begin{itemize}
  \item index of the closest cluster in the Milky Way Star Cluster catalog of
  \citet{Kharchenko:2016aa} using 3D separation
  \item distance to the closest cluster in pc
  \item ID of the OB association (A-L) if the star is in the catalog of
  stars in the known OB associations by \citet{de-Zeeuw:1999aa}
\end{itemize}
%
A part of the entire table is presented in Table~\ref{tab:catalog}.

@@

In this \documentname, we looked for co-moving pairs of stars in the
\tgas\ data released as part of the \gaia\ DR1.
Our method is to compare
the fully marginalized likelihoods between the two hypotheses, that a pair of
stars shares the same 3D velocity, and that the two stars have two independent
3D velocities, incorporating the covariances of parallax and proper motions.
We argued for a reasonable cut of the likelihood ratio, and found
13,058 candidate co-moving pairs among 10,606 stars
with separation ranging from 0.005~pc to 10~pc, the limit of our search.

@@

We have also found a surprisingly large number of very wide separation ($>3$~pc)
mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
of each other, and not part of large connected components.
These are most likely remnants of dissolving wide binaries.
If so, this population should still be relatively young compared
to the general disk field population. Modeling the color-magnitude diagram
distribution of these stars can shed some light on this issue.
If they are remnants of dissolving systems that were born coeval,
the sample of very wide separation co-moving pairs can potentially be used
to measure the recent ($\lesssim 1$~Gyr)
star formation history in the Solar neighborhood.
Co-moving stars with separation less than 1~pc are very promising candidates
for wide binaries. They are found to be pairs of stars of varying stellar types.
Some of these pairs, such as main sequence-red giant or F/G/K-M dwarfs,
will be particularly valuable for testing theoretical
stellar models and calibrating observational measurements of low mass stars.

We note that a similar search for wide binaries using the \tgas\ data is performed
in a recent work by \citet{2016arXiv161107883O}. They find $\approx 1900$ wide binaries
with separation typically less than 1.5~pc, and 256 pairs with separation larger than
$\sim 1$~pc.
We emphasize that our method is based on a probabilistic model for the assumptions
on the \emph{3D velocities} of the two stars in a pair, and that we
marginalize over the (unknown) distances and velocities of the stars
in contrast to just applying a cut in the proper motion space.
